### 2018-01-08 response_to_referee.txt

Dear Frank:

Please find enclosed our revised version of MS#AAS07769, Data analysis
recipes: Using Markov Chain Monte Carlo by Hogg & Foreman-Mackey.

In what follows, the referee comments are indented and our responses
are not.

Thank you,

David

     ---------------------------------------------------------------------- 
     Referee Report

     The authors provide a simple, and somehow anecdotal discussion
     about MCMC methods (for astronomers, although the title don't
     explicitly says it) in an informal and pragmatic style.

We figured that because this is in the AAS journals, that would supply
the relevant context, though we wouldn't object to modifying the title
slightly.

     The concept of conveying an intuitive description of statistical
     methods for the broad astronomical community is certainly
     something I support, however I would like to ask the authors and
     editor to consider some caveats below.

     Novelty and references

     The draft lacks of some important literature contextualization. A
     critical one is a recent Annual Review of Astronomy and
     Astrophysics entitled Markov Chain Monte Carlo Methods for
     Bayesian Data Analysis in Astronomy
     https://arxiv.org/abs/1706.01629

Yes, we know this paper. We now cite it [WHERE].

     I would like the authors to please explain how their review
     brings something new to the literature, considering the amount of
     good and available material about MCMC.

Good point. We are more explicit about this [WHERE]. The idea is to
delve deeper here, but on MCMC specifically, but not so much about new
developments; more like we are trying to specifically emphasize the
importance of basic good pratices; we say that more clearly now.

     Such as e.g.

     Doing Bayesian Data Analysis, Second Edition: A Tutorial with R,
     JAGS, and Stan, by John Kruschke, and many others

We also now cite some of this literature [WHERE; DO THIS].

     A potential reason is to update the readers about recent
     developments in the field.  But again, the authors seems to
     ignore the current astronomical literature. For example, the
     authors are aware of Statistical Bayesian packages such as JAGS
     and Stan, but do not cite two recent books in Astronomy that make
     use of them, such as:

     Bayesian Methods for the Physical Sciences: Learning from
     Examples in Astronomy and Physics, Springer Series in
     Astrostatistics, 2015 Andreon, and Weaver

     And

     Bayesian Models for Astrophysical Data: Using R, JAGS, Python,
     and Stan, Cambridge University press, Hilbe, de Souza and Ishida,
     2017

Good point; these are all now cited [WHERE].

     2. Confidence versus credible intervals

     I am sure the authors are aware of the difference between both,
     and should know they don't even agree beyond simple cases: such
     as for normally distributed data. Hence the quote "Because they
     are so afraid of being confused with frequentists, Bayesians
     often call these regions "credible" rather than "confidence""

     Not only seems inappropriate for a scientific paper, but it is
     actually misleading. See for instance The fallacy of placing
     confidence in confidence intervals, Morey et al. 2015
     https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4742505/

     Moreover, such discussion -- Bayesian vs Frequentist -- has far
     deeper roots. See e.g. Probability theory, the logic of science,
     E.T. Jaynes, Cambridge University Press 2003.

Actually, our problem is with the *names*. We agree with the referee
that these two things are importantly and substantially different! We
have fixed that all up in the text now [DO THIS].

     3. Lack of discussion about prediction vs credible intervals.

[UHHH SUP?]

     4. Model check

     Lack of contemporary discussion such as e.g. Gabry et
     al. Visualization in Bayesian workflow,
     https://arxiv.org/abs/1709.01449

Ooh that's a great reference! We cite this and were pleased to be
pointed at this.

     5. MORE SOPHISTICATED SAMPLING METHODS

     Please consider adding some material about

     Approximate Bayesian Computation, since it has gained space in astronomy e.g.

     https://arxiv.org/abs/1608.07606
     https://arxiv.org/abs/1607.01782 
     https://arxiv.org/abs/1504.06129 
     https://arxiv.org/abs/1202.1426

Good point -- we are even working on this ourselves. We added a small
bit of discussion about this [WHERE].

     ---------------------------------------------------------------------- 
     Statistics Editor Comments:

     As a didactic presentation, I suggest two additions:

     (a) Add a table listing MCMC approaches so the reader appreciates
     the variety of options available
     (e.g. https://www.rdocumentation.org/packages/LaplacesDemon/versions/16.0.1/topics/LaplacesDemon)

     (b) Discuss how to choose when MCMC sampling and when
     optimization (e.g. EM Algorithm or simplex) is appropriate. The
     former is commonly used for characterizing Bayesian posteriors
     while the latter are commonly used for maximizing likelihoods.

### 2018-01-21 mcmc.tex
no todos
commentstrip
flatten for submission
one change done

% to-do
% -----
% - Go through github issues, if any.
% - Post on arXiv
% - Submit to ApJ special issue
% - Blog about it

@@

\documentclass[modern]{aastex61}
\include{styles/dar}

@@

In what follows, we are going to provide a kind of ``user manual'' or
advice document or folklore capture regarding the use of \MCMC\ for
data analysis.
This will not be a detailed description of multiple \MCMC\ methods
(indeed, we will only explain one method in detail), and it will not
be about the mathematical properties or structure of the method or
methods.
It will be about how to use \MCMC, including diagnosis and
trouble-shooting.

@@

%Of course, the autocorrelation-time definition
%(\ref{eq:autocorrelationtime}) does not by itself provide an
%\emph{estimator} for the time. Estimating the auto-correlation time is
%notoriously difficult: It is a two-point statistic, and two-point statistics
%are much harder to estimate than one-point statistics.\note{See: All of cosmology!}
%If you are not a hard-core user of \MCMC, and if all you want is heuristic indicators
%of convergence, then what you should take from this section is that the auto-correlation
%time is involved in variance estimates through \equationname~(\ref{eq:tauvar}) but that
%it is hard to estimate.

%If you want to think about autocorrelation estimation:\note{One of us (DFM) has written
%  much more about autocorrelation estimation in the \project{emcee} documentation
%  at \url{http://dfm.io/posts/autocorr}.}
%You might first try the na\"ive estimator
%\begin{eqnarray}\label{eq:tauestnaive}
%\tau &\stackrel{?}{=}& 1 + 2\,\sum_{\Delta = 1}^{N} C(\Delta)
%\end{eqnarray}
%where the sum is over the autocorrelation function estimated based the chain
%using \equationname~(\ref{eq:autocorrelationfn}).
%In practice, this is a high variance estimator of the integrated
%autocorrelation time and this variance doesn't go to zero for an infinitely
%long chain.\note{%
%  A.~Sokal has a clear discussion of this topic and several proposals for
%  robust autocorrelation time estimation in a set of lecture notes
%  \url{http://www.stat.unc.edu/faculty/cji/Sokal.pdf}.}
%Instead, we can get a much more robust estimate of the autocorrelation time by
%restricting the sum in \equationname~(\ref{eq:tauestnaive}) to $\Delta \le M$
%where
%$\tau \le M \ll N$
%\begin{eqnarray}\label{eq:tauest}
%\tau &=& 1 + 2\,\sum_{\Delta = 1}^{M} C(\Delta)
%\end{eqnarray}
%In practice, $M$ can be chosen iteratively to satisfy these constraints but it
%will only be as reliable as the chain itself.
%An estimate of $\tau$ using \equationname~(\ref{eq:tauest}) should only be
%trusted if the chain is many times ($\sim 100$) longer than $\tau$.\footnote{This
%  point is also emphasized at the aforementioned tutorial
%  \url{http://dfm.io/posts/autocorr}, where
%  a number like 50 is recommended.}

Besides estimating the integrated autocorrelation time, another simple and
sensible test of convergence is the Gelman--Rubin
diagnostic\note{\citet{gelmanrubin}.}, which compares the variance (in one
parameter, or your most important parameter, or all parameters) \emph{within}
a chain to the variance \emph{across} chains.
This requires running multiple chains, and looking at the empirical
variance of the parameter away from its mean within each individual
chain, and comparing it to the variance in the mean of that parameter
across chains, inflated to be a mean per-sample variance.
What Gelman and Rubin do with these variances specifically is
sensible, but the important question of convergence is whether, as the
chains get longer, these two variances asymptotically reach stable
values, and that those two values agree.
The Gelman--Rubin diagnostic is related to the autocorrelation time,
in that it will only deliver success when the chains are much longer
than an autocorrelation time.
%It can also be used to assess the accuracy of the chain for performing
%integrals---by replacing the mean of the parameter with whatever
%integral is of most interest.

@@

For example,
  imagine that by \MCMC\ you have generated
  $K$ samples $\pars_k$ of a parameter vector $\pars$.
Now you have a scalar function $g(\pars)$ which takes the parameter vector
  and returns a scalar value%
  \note{For the median-of-sampling value, $g(\pars)$ needs to return a scalar,
  but technically, for the mean-of-sampling, $g(\pars)$ can be a vector or something high-dimensional.}
  which could be as simple as a single component of the parameter vector
  (one parameter from the list),
  or something more complex.
The mean-of-sampling value $\mean{q}$ is just
\begin{eqnarray}
\mean{q} &\leftarrow& \frac{1}{K}\,\sum_{k=1}^K g(\pars_k)
\quad ,
\end{eqnarray}
  and the median-of-sampling value is just the $[K/2]$th value of $g(\pars_k)$
  when the $g(\pars_k)$ have been ordered (sorted) from lowest to highest.
These are both produced by integrals;
  the first is the value returned by the expectation integral estimate,
  the second is the value past which half of the integrated pdf lies.

@@

%% ...Finally, to repeat: You can't sample the likelihood!  You can only
%% sample a posterior pdf!

@@

All we are going to do here is give a cursory mention to some methods
we know about, with some keywords that can be used to search more
deeply.
There is no sense in which this list is comprehensive; these are just
methods (or method classes) that are likely to be useful to projects similar
to projects we ourselves have executed.

@@

%\clearpage
\acknowledgements
We would like to thank
          Alex Barnett (Flatiron),
          Jo Bovy (Toronto),
          Brendon Brewer (Auckland),
          Will Farr (Flatiron),
          Marla Geha (Yale),
          Jonathan Goodman (NYU),
          Fengji Hou (NYU),
          Dustin Lang (Toronto),
          Phil Marshall (KIPAC), and
          Iain Murray (Edinburgh)
        for valuable advice and comments over the years, and
          Alessandro Gentilini
        for a very close read.
        This project benefitted from various research grants, especially
          NSF grants AST-0908357, IIS-1124794, and AST-1517237,
          NASA grant NNX12AI50G,
          and the The Moore-Sloan Data Science Environment at NYU.
This work was performed in part under contract with the Jet Propulsion
Laboratory (JPL) funded by NASA through the Sagan Fellowship Program executed
by the NASA Exoplanet Science Institute.

@@

>>>>

@@

\documentclass[modern]{aastex61}
\include{dar}

@@

In what follows, we are going to provide a kind of ``user manual'' or
advice document or folklore capture regarding the use of \MCMC\ for
data analysis.
This will not be a detailed description of multiple \MCMC\ methods
(indeed, we will only explain one method in detail), and it will not
be about the mathematical properties or structure of the method or
methods.\note{For a description of various different sampling methods,
  given from the point of view of an astrophysicist, see the review by
  \citet{sharma}. We say a bit about different \MCMC\ methods in
  \sectionname~\ref{sec:methods}, but our discussion is limited
  relative to Sharma's review.}
It will be about how to use \MCMC, including diagnosis and
trouble-shooting.

@@

Besides estimating the integrated autocorrelation time, another simple and
sensible test of convergence is the Gelman--Rubin
diagnostic\note{\citet{gelmanrubin}.}, which compares the variance (in one
parameter, or your most important parameter, or all parameters) \emph{within}
a chain to the variance \emph{across} chains.
This requires running multiple chains, and looking at the empirical
variance of the parameter away from its mean within each individual
chain, and comparing it to the variance in the mean of that parameter
across chains, inflated to be a mean per-sample variance.
What Gelman and Rubin do with these variances specifically is
sensible, but the important question of convergence is whether, as the
chains get longer, these two variances asymptotically reach stable
values, and that those two values agree.
The Gelman--Rubin diagnostic is related to the autocorrelation time,
in that it will only deliver success when the chains are much longer
than an autocorrelation time.

@@

For example,
  imagine that by \MCMC\ you have generated
  $K$ samples $\pars_k$ of a parameter vector $\pars$.
Now you have a scalar function $g(\pars)$ which takes the parameter vector
  and returns a scalar value%
  \note{For the median-of-sampling value, $g(\pars)$ needs to return a scalar,
  but technically, for the mean-of-sampling, $g(\pars)$ can be a vector or something high-dimensional.}
  which could be as simple as a single component of the parameter vector
  (one parameter from the list),
  or something more complex.
The mean-of-sampling value $\mean{g}$ is just
\begin{eqnarray}
\mean{g} &\leftarrow& \frac{1}{K}\,\sum_{k=1}^K g(\pars_k)
\quad ,
\end{eqnarray}
  and the median-of-sampling value is just the $[K/2]$th value of $g(\pars_k)$
  when the $g(\pars_k)$ have been ordered (sorted) from lowest to highest.
These are both produced by integrals;
  the first is the value returned by the expectation integral estimate,
  the second is the value past which half of the integrated pdf lies.

@@

@@

All we are going to do here is give a cursory mention to some methods
we know about, with some keywords that can be used to search more
deeply.
There is no sense in which this list is comprehensive; these are just
methods (or method classes) that are likely to be useful to projects
similar to projects we ourselves have executed.\note{There is a
  more in-depth and more comprehensive list of astrophyically relevant
  \MCMC\ methods in the review by \citet{sharma}.}

@@

\acknowledgements
We would like to thank
          Alex Barnett (Flatiron),
          Jo Bovy (Toronto),
          Brendon Brewer (Auckland),
          Will Farr (Flatiron),
          Marla Geha (Yale),
          Jonathan Goodman (NYU),
          Fengji Hou (NYU),
          Dustin Lang (Toronto),
          Phil Marshall (KIPAC), and
          Iain Murray (Edinburgh)
        for valuable advice and comments over the years, and
          Alessandro Gentilini
        for a very close read.
        This project benefitted from various research grants, especially
          NSF grants AST-0908357, IIS-1124794, and AST-1517237,
          NASA grant NNX12AI50G,
          and the The Moore-Sloan Data Science Environment at NYU.
This work was performed in part under contract with the Jet Propulsion
Laboratory (JPL) funded by NASA through the Sagan Fellowship Program executed
by the NASA Exoplanet Science Institute.

@@

\bibitem[Sharma(2017)]{sharma}
  Sharma, S., 2017,
  \href{http://adsabs.harvard.edu/abs/2017ARA&A..55..213S}{Markov Chain Monte Carlo methods for Bayesian data analysis in astronomy},
  \textit{Ann.\ Rev.\ Astron.\ Astrophys.} \textbf{55} 213
  \doi{10.1146/annurev-astro-082214-122339}.
